---
title: "Database_environmental_data"
output: html_document
date: "2024-11-26"
---

```{r}
library(tidyverse)
library(sf)

library(raster)
```


Dataset for extracting the environmental data for each sample of the database


Calculamos el numero Hill

```{r}
pmodern <- readRDS("Pmodern_revised.rds")
```


```{r}
pmp_meta <- pmodern$pmp_meta
pmp_pollen <- pmodern$pmp_pollen
```



#Calculation of the Hill number
```{r}
hill <- pmp_pollen |>
  dplyr::select(Abies:Zollernia) |>
  analogue::n2("sites") |>
  as_tibble() |>
  rename(hill_n2 = value)
  

pmp_ecological <- bind_cols(pmp_pollen, hill) |>
  select(ID, hill_n2)


ggplot2::ggplot(data=pmp_ecological, aes(x=hill_n2)) +
  geom_histogram(bins = 100, color="gray60", fill="gray60") +
  scale_x_continuous(n.breaks = 20, expand = c(0,0)) +
 # geom_vline(xintercept = 2, linetype="dotted",  color = "red", size=0.8) + 
  labs(title = "Hill number", subtitle = "Pmodern dataset") +
  theme_test()
```


Interest region
```{r}
latam_polygon <- rnaturalearth::ne_countries(scale = 'large', returnclass = "sf") |>
  dplyr::filter(name %in% c("Mexico", "Guatemala", "Belize", "El Salvador", "Honduras", "Nicaragua", "Costa Rica", "Panama", "Dominican Rep.","Cuba", "Colombia", "Venezuela", "Ecuador", "Guyana", "Suriname", "French Guiana", "Peru", "Brazil", "Argentina", "Bolivia", "Uruguay", "Paraguay", "Chile","Puerto Rico","Jamaica")) |>
  sf::st_union() |>
  sf::st_crop(ymin = -34, ymax = 22, xmin=-106, xmax=-35) |>
  sf::st_as_sf() 

sh_polygon <- rnaturalearth::ne_countries(scale = 'large', returnclass = "sf") %>% filter(name %in% c("Venezuela", "Colombia", "Guyana", "Suriname", "France", "Antigua and Barbuda", "Dominica", "Ecuador", "Peru", "Uruguay", "Bolivia", "Paraguay", "Argentina", "Chile", "Brazil")) |>
  sf::st_transform(crs = "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0 +units=m +no_defs") |>
  sf::st_crop(xmin = -85, xmax = -30, ymin = -60, ymax = 11) |> sf::st_union() |>
  sf::st_as_sf() 

nh_polygon <- rnaturalearth::ne_countries(scale = 'large', returnclass = "sf") %>%
  filter(name %in% c("Mexico","Belize","Guatemala","Honduras","El Salvador","Nicaragua","Costa Rica","Panama", "Venezuela", "Colombia", "Guyana", "Suriname", "France", "Antigua and Barbuda", "Dominica", "Ecuador","Brazil")) |>
  sf::st_transform(crs = "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0 +units=m +no_defs") |>
  sf::st_crop(xmin = -110, xmax = -50, ymin = 5, ymax = 22) |> sf::st_union() |>
  sf::st_as_sf()

ggplot2::ggplot() + ggplot2::geom_sf(data = sh_polygon) + ggplot2::theme_minimal()
ggplot2::ggplot() + ggplot2::geom_sf(data = nh_polygon) + ggplot2::theme_minimal()
ggplot2::ggplot() + ggplot2::geom_sf(data = latam_polygon) + ggplot2::theme_minimal()
```

# Global terrestrial human footprint

We start by the Global terrestrial human footprint
Venter et al. 2016: https://datadryad.org/stash/dataset/doi:10.5061/dryad.052q5

El raster original es de resolucion 1km (i. est. each pixel is 1km per side)
```{r, warning=FALSE}
input_file <- "Input data/HFP2009.tif"
output_file <- "Output_data/HFP2009_reprojected.tif"

# Reproject with gdalwarp
#gdalUtilities::gdalwarp(srcfile = input_file, dstfile = output_file, t_srs = "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0 +no_defs")
```



```{r}
raster_raw <- raster::raster(output_file)

footprint_raster <- raster_raw |>
  raster::crop(raster::extent(-108, -33, -40, 25)) |> 
  raster::mask(latam_polygon)


#footprint_tibble <- raster_raw |>
  #raster::crop(raster::extent(-108, -33, -40, 25)) |> 
  #raster::mask(latam_polygon) |>
  #raster::rasterToPoints() |>
  #as_tibble() |>
  #rename(Pressure = HFP2009_reprojected)
```


```{r}
pmp_meta$human_footprint_exact <- raster::extract(footprint_raster, pmp_meta[, c("lon", "lat")])
```



```{r}
# Step 1: Convert pmp_meta to an sf object
pmp_meta_sf <- st_as_sf(pmp_meta, coords = c("lon", "lat"), crs = 4326)

# Step 2: Create a 12 km buffer around each point (CRS 4326 uses degrees, so we will reproject to meters)
pmp_meta_sf <- st_transform(pmp_meta_sf, crs = 3395)  # Reproject to a CRS with meters (EPSG:3395)

# Create a 12 km buffer (12,000 meters)
pmp_meta_sf$buffer <- st_buffer(pmp_meta_sf, dist = 12000)

ggplot() +
  geom_sf(data = pmp_meta_sf$buffer, fill = "transparent", colour="blue") +  # Plot
  geom_sf(data=pmp_meta_sf$geometry, size = 0.5, colour="red") +
  labs(title = "Sample Points with 12 km Buffers") +
  theme_minimal()

# Step 3: Reproject the raster to the same CRS as the points (EPSG:3395)
footprint_raster_3395 <- projectRaster(footprint_raster, crs = CRS("+init=epsg:3395"))
#writeRaster(footprint_raster_3395, "Output_data/footprint_raster_3395.tif", format = "GTiff", overwrite = TRUE)


pmp_meta_sf$HFP2009_avg <- extract(footprint_raster_3395, pmp_meta_sf$buffer, 
                                    fun = function(x, ...) {
                                        if (length(x) > 0) {
                                            # Remove NA values before processing
                                            x <- na.omit(x)
                                            # Calculate the average (mean) of the values
                                            return(mean(x))
                                        } else {
                                            return(NA)
                                        }
                                    })


#saveRDS(pmp_meta_sf, "Output_data/pmp_meta_sf_average.rds")

pmp_meta_sf <- readRDS("Output_data/pmp_meta_sf_average.rds")

pmp_meta_sf <- st_transform(pmp_meta_sf, st_crs(latam_polygon))

# Step 2: Extract coordinates, drop geometry, and convert the buffer column to numeric
pmp_meta_df <- pmp_meta_sf %>%
  mutate(
    human_footprint_avg12km = as.numeric(HFP2009_avg), # Convert matrix to numeric
    longitude = st_coordinates(.)[, 1],   # Extract longitude
    latitude = st_coordinates(.)[, 2]     # Extract latitude
  ) %>%
  st_drop_geometry() %>%                  # Remove spatial geometry
  dplyr::select(-c(buffer, HFP2009_avg))


ggplot() + 
  geom_sf(data = latam_polygon) + 
  theme_minimal() +
  geom_point(data = pmp_meta_df, 
             aes(x = longitude, y = latitude, fill = human_footprint_avg12km, color = human_footprint_avg12km), size=0.5) +
  scale_color_viridis_c(option = "plasma") + # Optional: Adjust the color scale
  scale_fill_viridis_c(option = "plasma") 
```



```{r}
pmp_ecological <- inner_join(pmp_meta, pmp_meta_df) |>
  inner_join(pmp_ecological) |>
  dplyr::select(ID, hill_n2, human_footprint_exact, human_footprint_avg12km)
```

#=====
# Ahora obtenemos el indice de fragmentcion de la vegetacion


Global Forest Fragmentation 
Ma et al. 2023: https://figshare.com/s/21dbf1f50250aeb7f5a0 

El raster original es de resolucion 5km (i. est. each pixel is 5km per side)
```{r, warning=FALSE}
input_file <- "Input data/FFI2000.tif"
output_file <- "Output_data/FFI2000_reprojected.tif"

# Reproject with gdalwarp
#gdalUtilities::gdalwarp(srcfile = input_file, dstfile = output_file, t_srs = "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0 +no_defs")
```



```{r}
raster_raw <- raster::raster(output_file)

forest_raster <- raster_raw |>
  raster::crop(raster::extent(-108, -33, -40, 25)) |> 
  raster::mask(latam_polygon)

forest_raster

#forest_tibble <- raster_raw |>
  #raster::crop(raster::extent(-108, -33, -40, 25)) |> 
  #raster::mask(latam_polygon) |>
  #raster::rasterToPoints() |>
  #as_tibble() |>
  #rename( = HFP2009_reprojected)

```


```{r}
pmp_meta$forest_fragment_exact <- raster::extract(forest_raster, pmp_meta[, c("lon", "lat")])
```


```{r}
pmp_ecological <- left_join(pmp_meta, pmp_ecological) |>
  dplyr::select(ID, hill_n2, human_footprint_exact, human_footprint_avg12km, forest_fragment_exact)
```


```{r}
# Create a list containing the three objects
data_list <- list(pmp_meta = pmp_meta, pmp_pollen = pmp_pollen, pmp_ecological = pmp_ecological)

# Save the list as an .rds file
saveRDS(data_list, file = "Pmodern_revised.rds")

```







```{r}
plot_hfp <- ggplot() + 
  geom_sf(data = latam_polygon, fill="transparent") +
  geom_tile(data = footprint, aes(x = x, y = y, fill = Pressure)) + 
  scico::scale_fill_scico(palette = 'lipari') +
  scico::scale_colour_scico(palette = 'lipari') +
  geom_point(data = pmp_meta, aes(x=lon, y=lat), shape =21, fill="white", colour="gray20", stroke = 0.1, size=0.45) +
  labs(subtitle = "Terrestrial human footprint (Venter et al. 2016)") +
  scale_x_continuous(expand = c(0.015,0.015)) +
  scale_y_continuous(expand = c(0.015,0.015)) +
  ggspatial::annotation_north_arrow(location = "tr", which_north = "true",height = unit(4.8, "mm"), width = unit(4.8, "mm"), pad_x = unit(2, "mm"), pad_y = unit(2, "mm"),style = ggspatial::north_arrow_nautical(fill = c("grey40", "white"),line_col = "grey20", text_family = "Book Antiqua",text_size = 5.5)) #+  # "Mongolian Baiti"
  # theme(plot.subtitle = element_text(size = 7.5),
        # panel.background = element_rect(fill="transparent",color="transparent"),
        # panel.border = element_rect(fill="transparent",color="black"),
       # panel.grid = element_blank(),
       # axis.title.x = element_text(size=7.5,color = "black"),
       # axis.title.y = element_text(size=7.5, angle=90),
       # axis.text.x = element_text(size=7.5),
       # axis.text.y = element_text(size=7.5),
       # legend.key.size = unit(0.28,"cm"),
       # legend.text.align = 0,
       # legend.text=element_text(size=7.5),
       # legend.title.align = 0.5,
       # legend.title = element_text(size = 7.5),
       # legend.margin = margin(unit = "cm", t=0.05,r=0.08,b=0.1,l=0.08),
       # legend.box.background = element_rect(fill=alpha('white', 0.2),color="gray30",size = 0.15,),
       # legend.background = element_rect(fill=alpha('white', 0.45),color="gray30",size = 0.25),
       # legend.position = c(0.13, 0.13),
       # plot.margin = margin(0, 0, 0, 0)) 

plot_hfp
```


```{r}
ggsave(plot = plot_hfp, "Output_figures/Human_footprint.jpg", 
       units = "cm", width = 10, height = 10, dpi = 700, quality = 100)
```



Here it is we obtain the human impact for each sample without using a buffer
```{r}
# Convert pmp_meta and footprint to sf objects (coordinates in WGS84: EPSG:4326)
pmp_meta_sf <- sf::st_as_sf(pmp_meta, coords = c("lon", "lat"), crs = 4326)
footprint_sf <- sf::st_as_sf(footprint, coords = c("x", "y"), crs = 4326)

# Perform a spatial join to find the nearest pressure value for each point in pmp_meta
# st_nearest_join finds the nearest feature
joined_sf <- sf::st_nearest_join(pmp_meta_sf, footprint_sf, by = c("lon" = "x", "lat" = "y"))

# Extract the joined result, including "Pressure"
result_df <- as_tibble(joined_sf) %>%
  select(ID, sample, sourcetype, lon, lat, Elevation, Pressure)

# View the result
print(result_df)
```





```{r}
sample_n(pmp_meta, size = 3)

sample_n(footprint, size = 3)
```









```{r}
maps <- sf::st_read("C:/Users/esmer/UoR/latam_climate_reconstruction/Input/Precipitation_polygon/my_new_polygon.shp") |> 
  filter(Value == 17)
p2_EPA_north_andes <- sf::st_intersection(maps, sh_polygon) |>
  filter(Value == "17")

p0_andes <- sf::st_read("C:/Users/esmer/UoR/latam_climate_reconstruction/Output_data/andes_level_17_18.shp")
p3_north_andes <- sf::st_read("C:/Users/esmer/UoR/latam_climate_reconstruction/Output_data/north_andes.shp")
p1_north_hemis <- sf::st_read("C:/Users/esmer/UoR/latam_climate_reconstruction/Output_data/mean_summer_regions_sf_nh_1.shp")
p4_sasm2 <- sf::st_read("C:/Users/esmer/UoR/latam_climate_reconstruction/Output_data/sasm_core2.shp")
p5_sasm3 <- sf::st_read("C:/Users/esmer/UoR/latam_climate_reconstruction/Output_data/sasm_core3.shp")
p6_sasm4 <- sf::st_read("C:/Users/esmer/UoR/latam_climate_reconstruction/Output_data/sasm_core2_noITCZ_3.shp")
p6_sasm4 <- sf::st_make_valid(p6_sasm4)


ggplot2::ggplot() + 
  ggplot2::geom_sf(data = latam_polygon) + 
  geom_sf(data = p2_EPA_north_andes, colour="pink", fill="pink", alpha = 0.5) +
  geom_sf(data = p0_andes, colour="skyblue1", fill="skyblue1", alpha = 0.4) +
  geom_sf(data = p3_north_andes, colour = "palegreen", fill= "palegreen", alpha = 0.3) +
  geom_sf(data = p1_north_hemis, colour = "orchid1", fill= "orchid1", alpha = 0.1) +
 # geom_sf(data = p5_sasm3, colour="salmon", fill="salmon", alpha = 0.3) +
 
  # geom_sf(data = p6_sasm4, colour="red", fill="red", alpha  = 0.1) +
   # geom_sf(data = p4_sasm2, colour="khaki1", fill="khaki1", alpha = 0.3) +
  ggplot2::theme_minimal()
```


```{r}
maps <- sf::st_read("C:/Users/esmer/UoR/latam_climate_reconstruction/Input/Precipitation_polygon/my_new_polygon.shp") 
maps1 <- sf::st_intersection(maps, nh_polygon) 

ggplot2::ggplot() + 
  ggplot2::geom_sf(data = nh_polygon) +  
  geom_sf(data = filter(maps1, !Value%in%c(0,10,9)), aes(fill=Value, colour=Value), alpha = 0.3) +
 # ggplot2::geom_point(data = filter(pmp_meta, lat > 5), aes(x=lon, y=lat), size=0.3) +
 # geom_hline(yintercept=18, linetype="dashed", color = "red") +
  #geom_vline(xintercept=-95, linetype="dashed", color = "red") +
  #geom_vline(xintercept=-91.85, linetype="dashed", color = "blue")
```





```{r}
pmp_meta

pmp_pollen

pmp_ecological
```








